### 2025-02-18 pre_print/pre_print.qmd
DOI added to pre print

---
articletitle: |
  The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members
# shorttitle: "Doe et al. (2022)" # usually auto-generated
format: 
  sportrxiv-pdf:
    include-in-header:
      text: |
       \usepackage[font=scriptsize]{caption}
       \usepackage{caption}
author:
  - name: James Steele
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james@steele-research.com
  - name: James Fisher
    affiliations:
      - ref: 3
      - ref: 4
    orcid: 0000-0002-6013-8402
    email: james.fisher.phd@outlook.com
  - name: Tim Dettmann
    affiliations:
      - ref: 1
    email: tim.dettmann@gmail.com
affiliations:
      - id: 1
        name: Kieser Australia, Melbourne, Australia
      - id: 2
        name: Steele Research Limited, Eastleigh, UK
      - id: 3
        name: Department of Sport and Health, Solent University, Southampton, UK
      - id: 4
        name: The Exercise Coach, Tennessee, USA
abstract: |
  Supervision during resistance training (RT) may enhance strength gains by optimizing trainee effort. We investigated supervision’s role in effort during RT in a unique setting with private strength clinics, where members train either unsupervised (“Core” membership) or supervised by a qualified exercise scientist (“Assisted” membership). Using both retrospective analysis of member training records and a prospective experimental study, we examined supervision’s impact on exercise performance, measured as time under load (TUL), rating of perceived effort (RPE), and rating of perceived discomfort (RPD). Bayesian methods were applied, using empirically informed prior distributions from retrospective data to model the experimental study. The prior sample included 1000 training sessions from each membership type, while the experimental study involved 45 Core members performing both supervised and unsupervised sessions in randomized order, using their current training loads to momentary failure. Our findings suggest that, in real-world settings (in situ), exercise performance differed little between supervised and unsupervised training. However, in our experimental study, supervision significantly improved TUL (Core = 125.12 [95%QI: 113.70, 131.90] sec; Assisted = 147.35 [95%QI: 134.29, 154.81] sec; contrast = -22.10 [95%QI: -26.60, -17.61] sec). RPE was slightly higher with supervision in both real-world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; contrast = -6% [95%QI: -8%, -4%]) and experimental settings (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; contrast = -6% [95%QI: -10%, -4%]), suggesting trainees push closer to failure under supervision. This was further supported by higher RPD during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3]; Assisted = 7.5 [95%QI: 6.5, 8.3]; contrast = -1.2 [95%QI: -1.6, -0.9]). Overall, these results reinforce prior research on the benefits of supervision in RT, indicating that unsupervised trainees—especially in real-world conditions—likely train with suboptimal effort.
license-type: ccby # change if neccessary
# year: 2025 # defaults to current year
keywords: [template, demo, exercise science] # optional
# optional link to Supplementary Material:
suppl-link: https://osf.io/etb34/
reference-section-title: References
printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
bibliography: bibliography.bib  
pdf-engine: xelatex
execute: 
  echo: false
  message: false
  warning: false
---

>>>>

---
articletitle: |
  The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members
# shorttitle: "Doe et al. (2022)" # usually auto-generated
format: 
  sportrxiv-pdf:
    include-in-header:
      text: |
       \usepackage[font=scriptsize]{caption}
       \usepackage{caption}
author:
  - name: James Steele
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james@steele-research.com
  - name: James Fisher
    affiliations:
      - ref: 3
      - ref: 4
    orcid: 0000-0002-6013-8402
    email: james.fisher.phd@outlook.com
  - name: Tim Dettmann
    affiliations:
      - ref: 1
    email: tim.dettmann@gmail.com
affiliations:
      - id: 1
        name: Kieser Australia, Melbourne, Australia
      - id: 2
        name: Steele Research Limited, Eastleigh, UK
      - id: 3
        name: Department of Sport and Health, Solent University, Southampton, UK
      - id: 4
        name: The Exercise Coach, Tennessee, USA
abstract: |
  Supervision during resistance training (RT) may enhance strength gains by optimizing trainee effort. We investigated supervision’s role in effort during RT in a unique setting with private strength clinics, where members train either unsupervised (“Core” membership) or supervised by a qualified exercise scientist (“Assisted” membership). Using both retrospective analysis of member training records and a prospective experimental study, we examined supervision’s impact on exercise performance, measured as time under load (TUL), rating of perceived effort (RPE), and rating of perceived discomfort (RPD). Bayesian methods were applied, using empirically informed prior distributions from retrospective data to model the experimental study. The prior sample included ~1000 members training sessions from each membership type, while the experimental study involved 45 Core members performing both supervised and unsupervised sessions in randomized order, using their current training loads to momentary failure. Our findings suggest that, in real-world settings (in situ), exercise performance differed little between supervised and unsupervised training. However, in our experimental study, supervision improved TUL (Core = 125.12 [95%QI: 113.70, 131.90] sec; Assisted = 147.35 [95%QI: 134.29, 154.81] sec; contrast = -22.10 [95%QI: -26.60, -17.61] sec). In percentage points RPE was slightly higher with supervision in both prior real-world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; contrast = -6% [95%QI: -8%, -4%]) and experimental settings (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; contrast = -6% [95%QI: -10%, -4%]), suggesting trainees push closer to failure under supervision. This was further supported by higher RPD during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3]; Assisted = 7.5 [95%QI: 6.5, 8.3]; contrast = -1.2 [95%QI: -1.6, -0.9]). Overall, these results reinforce prior research on the benefits of supervision in RT, indicating that unsupervised trainees—especially in real-world conditions—likely train with suboptimal effort.
license-type: ccby # change if neccessary
# year: 2025 # defaults to current year
keywords: [supervision, effort, resistance training, bayesian] # optional
# optional link to Supplementary Material:
suppl-link: https://osf.io/etb34/
reference-section-title: References
printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
bibliography: bibliography.bib  
pdf-engine: xelatex
execute: 
  echo: false
  message: false
  warning: false
---
